---
title: "Analyzing Movie Ratings"
author: "Abe Ceasar Perez"
date: "April 10, 2022"
output: html_document
---

### Extracting Data from a Webpage

This project aims to apply the practice of web-scraping using R to extract relevant data. Specifically, the goal is to extract the top 30 movies in the IMDB website from March to July 2020.

In addition to the objectives above, we'll be applying the following concepts in this project:

- Webscraping functions
- Vectorization
- String Manipulation
- Plotting/Visualization

To understand how the results were obtained, the process is outlined below:

### Loading the Packages

The following packages to be loaded are outlined here along with their uses:

- rvest - for loading the web-scraping functions
- stringr - for extracting/manipulating strings
- readr - for reading files/manipulating data
- dplyr - for loading additional functions
- ggplot2 - for visualizing data

```{r load_packages, echo=TRUE, warning=FALSE}

pacman::p_load("rvest", "stringr","dplyr", "ggplot2", "readr") # reads all package names and loads them simultaneously

```

### Loading the Webpage

We'll now start by reading the webpage content using the read_html function. the link that we'll be using is a copy of the page of interest to avoid external server instability issues.

```{r load_webpage, echo=TRUE, warning=FALSE}

url <- "http://dataquestio.github.io/web-scraping-pages/IMDb-DQgp.html"
wp_content <- read_html(url) # reads the link of the webpage
wp_content

```

### Extracting the Movies' Information

After we have read the html page, we'll now start extracing the movie ratings data from the read page. We'll begin by extracting the titles of the movies and storing them into a vector:

```{r extract_titles, echo=TRUE, warning=FALSE}

titles <- wp_content %>% 
  html_nodes(".lister-item-header a") %>% # reads the html tag for the movie title from the web page
  html_text() # converts the output into text

titles

```

Next, we'll be extracting the movies' year of releases. Since the years are wrapped in parenthesis, we'll be modfiying the string and then convert it into numerical form.

```{r extract_years, echo=TRUE, warning=FALSE}

years <- wp_content %>% 
  html_nodes(".text-muted.unbold") %>% # reads the html tag for the movie release year from the web page
  html_text() %>% 
  parse_number # extracts only the numerical component and changes the data type to integer

years

```

Using the similar process above, we'll also use that in extracting and modifying the runtime of the movies

```{r extract_runtime, echo=TRUE, warning=FALSE}

runtime <- wp_content %>% 
  html_nodes(".runtime") %>% # reads the html tag for movie runtime from the web page
  html_text() %>% 
  parse_number() # extracts only the numerical component and changes the data type to integer

runtime

```

For the movie genres, we'll also be extracting these similar to how the movie titles were extracted except that additional cleaning is made to remove the additional whitespace at the end of the genres.

```{r extract_genre, echo=TRUE, warning=FALSE}

genre <- wp_content %>% 
  html_nodes(".genre") %>% # reads the html tag for movie genre from the web page
  html_text() %>% 
  str_trim() # removes additional whitespace at the end of the string

genre

```

For the movie ratings, we'll be extracting it similar with the extracted numerical data above but using a different html function (html_attr). Apart from cleaning the data, we'll also be removing missing ratings data before storing it in a vector.

```{r extract_ratings, echo=TRUE, warning=FALSE}

ratings <- wp_content %>% 
  html_nodes("div") %>% 
  html_attr("data-value") # reads the corresponding html tags for movie ratings

ratings <- parse_number(ratings[!(is.na(ratings))]) # removes missing ratings before modifying the movie ratings

ratings

```

We'll also use this process to extract the metascores of each movie.

```{r extract_metascore, echo=TRUE, warning=FALSE}

metascore <- wp_content %>% 
  html_nodes("span.metascore") %>% 
  html_text() # reads the corresponding html tags for metascores

metascore <- parse_number(metascore) # extracts only the numerical component and changes the data type to integer

metascore

```

Similar to out previous process of extracting numerical data, we'll also be performing this in extracting the total votes obtained in the movies.

```{r extract_votes, echo=TRUE, warning=FALSE}

votes <- wp_content %>% 
  html_nodes(".text-muted+ span") %>% # reads the corresponding html tags for total movie votes
  html_text() %>% 
  parse_number() # extracts only the numerical component and changes the data type to integer

votes

```

### Modifying the Movie Vectors

Before combining all vectors into one dataframe, we'll be performing some vectorization to ensure that each vector is of the same length. 

For the first vector (metascore), the length of the vector (25) is short compared to the other vectors (titles, etc. - 30) so we'll be imputing values here using a user-defined function.

```{r impute_values, echo=TRUE, warning=FALSE}

append_vector <- function(vector, inserted_indices, values){

  ## Creating the current indices of the vector
  vector_current_indices <- 1:length(vector)

  ## Adding small amount of values (between 0 and 0.9) to the `inserted_indices`
  new_inserted_indices <- inserted_indices + seq(0, 0.9, length.out = length(inserted_indices))

  ## Appending the `new_inserted_indices` to the current vector indices
  indices <- c(vector_current_indices, new_inserted_indices)

  ## Ordering the indices
  ordered_indices <- order(indices)

  ## Appending the new value to the existing vector
  new_vector <- c(vector, values)

  ## Ordering the new vector wrt the ordered indices
  new_vector[ordered_indices]
}

new_metascores <- append_vector(metascore, c(1,1,1,13,24), NA) # calls the function along with imputed metascores

```

Next, since one movie title (A Quiet Place II) does not have majority of the information in each of the numerical vectors (rating, metascore, votes), we'll be removing all movie-related information for this title.

```{r remove_one_title, echo=TRUE, warning=FALSE}

quiet_place <- match("A Quiet Place Part II", titles)

titles <- titles[-quiet_place]
years <- years[-quiet_place]
runtime <- runtime[-quiet_place]
genre <- genre[-quiet_place]
new_metascores <- new_metascores[-quiet_place]

```

Combining all Vectors into one Dataframe

To summarize all the information from the vectors, we'll be combining all of them into one tibble/dataframe. We'll also be modifying the ratings into a single digit.

```{r dataframe, echo=TRUE, warning=FALSE}

df <- tibble(titles, 
             years,
             runtime, 
             genre, 
             ratings = floor(ratings), # keeps only the whole number of the rating
             new_metascores, 
             votes)
df

```

Lastly, to further understand the data in terms of movie rating, we'll be visualizing them using a boxplot.

```{r boxplot, echo=TRUE, warning=FALSE}

df %>% 
  ggplot(aes(x=as.factor(ratings), y=votes)) + # converts ratings into factor for categorization
  geom_boxplot() + 
  labs(x="rating")

```

<br>
<br>